### 2017-02-27 build/paper.tex
pdf/tex build

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Browsernative Microservices},
            pdfauthor={Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Browsernative Microservices}
\providecommand{\subtitle}[1]{}
\subtitle{Modular web architecture through new W3C specifications}
\author{Jan Peteler, FH Würzburg-Schweinfurt, jan.peteler@student.fhws.de}
\date{Januar 2017}

\begin{document}
\maketitle
\begin{abstract}
Building complex web applications nowadays require additional layers of
abstraction and often heavily depend on proprietary frameworks. New
specifications build right into the browserengine provide a native
service API to overcome tricky abstraction constraints.
\end{abstract}

{
\setcounter{tocdepth}{3}
\tableofcontents
}
\section{1. Simplicity and the web}\label{simplicity-and-the-web}

\begin{quote}
Simplicity is prerequisite for reliability. - Edsger W. Dijkstra
\end{quote}

Computers can scale, humans can't. Ever since a complex system made by
humans has been constrained by humans mental capabilities. Like in the
analogy of juggling balls our brain can just ``juggle'' a few things at
a time. Rich Hickey, the inventor of the programming language Clojure
gave an inspirational keynote on the topic of
\textbf{simplicity}.\footnote{\href{https://www.youtube.com/watch?v=rI8tNMsozo0\&t=46s}{Rails
  Conf 2012 Keynote: Simplicity Matters by Rich Hickey}} In every sphere
of a humans life simplicity aligns perception with our mental
capacities.

Derived from the Latin word \textbf{simplex}, simple can be understood
as ``literally, uncompounded or onefold''\footnote{\href{http://www.etymonline.com/index.php?term=simple}{Etymology
  Dictionary}} which points towards an unidimensional nature. While
complexity describes the multilayered und entangled nature of
conditions, simplicity empowers the human brain to reason about issues
in a straightforward manner. It certainly has some overlapping's with
easy, but while easy is more of a relative spirit, simple can be laid
out as a objective manner and therefore universally applicable.

Software development is undoubtedly rich in complexity and full of
subtle pitfalls. In a typical scenario a growing software project
evolves in one or another opinionated direction over time. Layers of
abstractions wrestling with aged legacy code requiring layers of
middleware. Mutating assets create subtle bugs and so forth. Eventually
the small piece of software may end up in a highly complected monolith
which will determine future design decisions to a painful degree.
Opaqueness of the system will slow down innovation to a minimum in the
need of ``keeping the lights on''.

On the other side of this dystopian scenario a truly modular system
architecture abandons many of those potential inconsistencies. The whole
system is divided in pluggable parts, object mutation is either
traceable or avoided altogether in favor of immutable data structures.
As Rich Hickey argues, design decisions should be made under the
\textbf{impression of extending, substitution, moving, combining and
repurposing}. The ability to reason about the program at any given time
is crucial for future decisions and implementations. Recalling again the
unidimensional nature of simplicity.

Simplicity in ``the web'', read as a loose generalization of
``everything that runs in the browser'' is certainly a story full of
misconceptions. While simplicity in the backend is mostly a matter of
principles and patterns, any browser-based frontend is restricted on the
highly deterministic nature of the browser platform.

In the last four years the average transfer size of a webpage doubled to
currently around 2.5 MB.\footnote{\href{http://httparchive.org/trends.php}{HTTPArchive
  Trends}} Leaving images, fonts or other content aside the size of
HTML, CSS and JS sums up to a total average of 550 kb. One character
weights around 1 byte which means an average webpage is delivering
550.000 character or around 125 pages of single-spaced text. Frederic
Filloux analyzed the payload on different newspaper websites and came to
the conclusion, that only round about 5-6 \% of the transferred
characters made for human consumption.{[}1{]}

Having an 95 \% overhead is rather undesirable for both the consumers
and creators of the website. Since it's a widespread problem without a
single point of failure one can argue the platform itself is the
failure. By design, every pageload results in a monolithic DOM tree
managed by the browser engine. Whether rendering just a bunch of static
text nodes or an ever changing webapp the underlying global nature of
the DOM tree remains the same. Every additional piece of code added to
the webpage will invisibly add another fold of complexity to this global
object.

In an non-deterministic runtime environment, encapsulation and
modularization is a typical pattern to make complexity manageable and
accommodate future uncertainty.{[}2, p. 1{]} Since years the average JS
payload is steadily rising which can be interpreted as a trend towards
more dynamic websites. The demands to the browser platform changed from
a static page renderer to a \textbf{dynamic UI machine} without changing
the underlying architecture significantly. Under the current situation
only additional layers of abstraction can wrestle complexity.

In the recent years many \textbf{frameworks}, libraries and
methodologies approached the global nature of the DOM by scoping assets
and design rules into maintainable components. While the DOM can't be
scoped, JS can. Many frameworks, like ReactJS, AngularJS or VueJS just
to name a few, ditched the old rule of separated HTML, CSS and JS in
favor of an additional layer of abstracted JS components (containing
structure, behavior and sometimes styling). Quiet often those frameworks
mimic a MVC pattern on top of the browser engine which is a reasonable
simple design pattern to build graphical user interfaces. While
frameworks are a valid approach for building scalable web applications
they remain highly opinionated, embody inherent complexity themselves
and can change and break over time. Another downside is code inflation
which is a crucial point for performance. All of those bottlenecks in
the web demand for a new standardized way for creating and evolving
complex web services.

In the year 2013 thinkers, creators and browser vendors joined together
to propose \emph{The Extensible Web Manifesto}.\footnote{\href{https://extensiblewebmanifesto.org/}{The
  Extensible Web Manifesto}} The manifestos claim was to enhance the
current web platforms with new low-level capabilities. Those features
should empower creators of the web to write more declarative code and
abandon known problems and artificial abstractions. Four years later,
the enhancement of JavaScript leapfrogged and many new low-level APIs
brought to life. With this new APIs at hand a vivid web developer can
create robust websites with less code and less additional libraries.
This paper is an approach to unfold these \textbf{browsernative}
technologies to create overall simple and resilient
\textbf{microservices} for the browser

\section{2. Microservices}\label{microservices}

In search of a better, simpler web architecture one might look on
already established pattern that proofed to fulfill enterprise needs.
Microservices are a good approach for tearing big monolithic systems
into fine-grained simple services with explicit defined boundaries. In a
nutshell a microservice is a small, autonomous service that works
together with other services seamlessly.{[}3, p. 2{]} Or with the words
of Fowler and Lewis: ``\ldots{} the microservice architectural style is
an approach to developing a single application as a suite of small
services, each running in its own process and communicating with
lightweight mechanisms, often an HTTP resource API.''{[}4{]} Yet at this
point a reader might spot some similarities with microservices and the
browser-based development: Both wrestling the problem of monolithic
architecture and both using lightweight communication mechanisms. In
fact, many big companies of ``the web'' like Amazon or Netflix
successfully transformed their monolithic system into a service based
system which gives a taste of the power behind microservices.{[}4{]}

Microservices incorporate a wide array of ideas from developing scalable
software, like \textbf{domain-driven design} where it pursues the
incorporation of real world structure in the code.{[}3, p. 2{]} Or
making use of \textbf{continuous delivery} for pushing software rapidly
through \textbf{automated deployment} mechanisms into production.{[}4{]}
Furthermore, microservices transcendent the technical perspective and
reaches into the team organization. As a primary source of truth this
paper relies on the work of Sam Newman, who has written a comprehensive
guide towards microservices{[}3{]} and the work of Fowler and
Lewis{[}4{]}. The purpose of this section is to gain confidence about
the microservices architecture in the context of the browser platform.

\subsection{2.1. Componentization via
Services}\label{componentization-via-services}

``A~\textbf{component}~is a unit of software that is independently
replaceable and upgradeable.'' {[}4{]} Components are the building
blocks of microservices. And microservices are the building blocks of
applications. Essentially the difference between microservices and
components is just the level of abstraction. Whether a concrete
microservice or a much more generic component, both share a similar set
of principles. Therefore this paper referring to both parts when talking
about \textbf{services}.

The first principle of services is the \textbf{loose coupling
principle}: changing and deploying one service shouldn't result in
changing other parts of the system.{[}3, p. 30{]}. Mutations or shadowed
variables, which is happening a lot in CSS, making it hard to keep
changes ought to only affect one place in the application. A
\emph{browsernative microservice} therefore pushing encapsulation and
avoiding variable mutations outside its scope as much as possible.
Practically, CSS will be scoped and JS fosters immutable JS entities and
avoids variables leaking into the global namespace.

The second principle of services is the \textbf{high cohesion
principle}: Whether designing a microservice or it's components we want
related behavior sit together, and unrelated behavior to sit
elsewhere.{[}3, p. 30{]} High cohesion can be enhanced towards the more
dynamic \textbf{Single Responsibility Principle}: ``Gather together
those things that change for the same reason and separate those things
that change for different reasons.''{[}5{]} In a very quick and dirty
code quality analysis, the quality can be measured just by counting the
places changes in the code occur in order to implement a functionality.
An arbitrary threefold MVC system should require a maximum of three
changes to implement or change functionalities. The problem in browser
based development is not only the global paradigm which makes changes
deliberately unpredictable. The high cohesion principle is violated by
the traditional separation along the siloed entities \textbf{HTML, JS
and CSS}. Understanding the relation of HTML markup and another CSS file
adds incidental complexity. Different approaches emerged over the recent
years to join the forces. Most notable the dedicated frontend language
\emph{elm}\footnote{\href{http://elm-lang.org/}{elm lang}} where MODEL,
VIEW and CONTROLLER sit in one place. A browsernative service aims for a
combinement of the web native trinity HTML, JS and CSS.

Traditional web development relies on \textbf{libraries} to enhance the
service capabilities of the web platform. Compared to libraries, a
component service offers multiple advantages for building, deployment
and shipping. As expressed earlier a \emph{components for the web}, or
web component, is \textbf{self-contained} which means it embodies all
needed functionality to get it's job done. Therefore it has a much
better evolution mechanism in the service contracts. Changing
functionality won't break other services. A component can progressively
enhanced which guarantees functionality throughout different versions. A
library is only loosely coupled to the implementation and therefore hard
to track in functionality. Changing a library may result in an
unforeseen amount of time fixing implementations. It is not unusual to
see websites embodying different versions of the same library to
guarantee functionality but lowers page performance significantly.

Another issue where web components stand out is related to performance
and especially the critical first page rendering. Libraries for the
browsers are traditionally ``shipped'' as non static immediately-invoked
functions run immediately on page load. Following the \emph{Google RAIL}
model, a user-centric performance measurement, a page load ought be less
than 1 second to catch up user's attention.{[}6{]} There are many ways
for optimizing the critical first render but as a rule of thumb a
build-in web component might be always superior to libraries in terms of
first rendering. The fine-grained lifecycle methods which will be
described in the technical section of this paper give the developer far
reaching optimization opportunities.

The last argument in favor of components over libraries is the more
explicit interface.{[}4{]} While the functionality of a library needs
documentation to be accessible a component functionality is exposed via
the components' signature. The markup language HTML is by design
equipped with an expressive syntax and steering a web component using
attributes and values should be most straightforward even to
unexperienced web developers.

\subsection{2.2. Organized around Business
Capabilities}\label{organized-around-business-capabilities}

\begin{quote}
``organizations which design systems \ldots{} are constrained to produce
designs which are copies of the communication~structures~of these
organizations''. {[}7{]}
\end{quote}

Emphasizing the human factor in microservices is a key feature.
Microservices are a product of real-world usage.{[}3, p. 1{]} Instead of
splitting team structures along the technology stack (UI Experts
-\textgreater{} Middleware -\textgreater{} Database) a microservice
approach model teams around \textbf{business capabilities}.{[}4{]}
Consequently every team is capable of planning, designing, implementing,
testing and maintaining their very own microservice. Along the
technology stack every member gains high competence about the the
service architecture which can be critical for evolving the service over
its whole lifecycle.

Real-world domains tend to be complex and multifaceted. To unfold their
complexity, domains can subdivided into \textbf{bounded context}.{[}3,
p. 31{]} For example, customer service is a business domain but with
varying bounded context. One context can be sales, another context could
be support. Every context makes different assumptions about the
underlying model and draws an explicit interface where it decides what
to share with other contexts.{[}3, p. 30{]} Evaluating each bounded
context within each business domain will eventually shape the data
persistency model likewise the interface to the service. This
methodology can be iterated over and over again. A sales service for
example might be evaluated to different sales context resulting in
differing interfaces on differing devices shaped by browsernative
microservices. By focusing the service into clear defined business
boundaries, it is easier to define a smart API of the service.

Assigning service responsibility to a team, the so called
\textbf{Definition of Done} (DoD) shifts from ``accomplishing projects''
to ``accomplishing products''. This new paradigm not only changes the
administrative overhead like budgeting or resource allocation. It
creates a kind of responsibility connection from the team to the service
which can be best described as \textbf{shared governance} model. ``Each
team collectively share responsibility for evolving the technical vision
of the system.''{[}3, p. 247{]} Expectedly, those teams are more
motivated within their very own service and exhibit a more sophisticated
iteration time.{[}4{]}

For many companies working in the spheres of the internet the client
side is highly important for their business. In fact, business goals and
capabilities can be derived from frontend needs. The state of the web is
not only a story of numerous artifacts, it is also a story of an highly
fragmented market along devices, operating systems, differing sizes and
functionalities. Different devices again have different assumptions
about the technology stack. Splitting teams along the stack results in
an slow paced back and forth negotiation for every change to be made
make iteration time expensive. As browser technologies, design
guidelines and devices change frequently it makes absolutely sense to
shift responsibility towards the teams altogether.

\subsection{2.3. Smart endpoints and dumb
pipes}\label{smart-endpoints-and-dumb-pipes}

To ensure the microservice functionality among teams and different
services requires thoughtful decentralization. Emphasizing once more the
real-world capabilities of microservices a message channel architecture
can be derived from patterns known from traditional postal services. A
physical letter has only two smart endpoints entitled to read and
process the message while packaging likewise the connection itself is
maximized towards unification. \emph{Smart endpoints and dumb pipes} is
coined to the approach of designing communication mostly decoupled and
as cohesive as possible.{[}4{]} Applying this rules to the web platform
can result into building unified JSON message objects passed along
``dumb'' middleware components. A typical browser event comes close to
this definition and suits arguably well for in-memory communication
between web components and microservices.

Microservices heavily rely on simple HTTP request-response with resource
APIs and lightweight messaging.{[}4{]} Newman puts his recommendation on
technologic-agnostic REST APIs to free data persistence from
implementation constraints.{[}3, p. 247{]} The advantage of this overall
simple communication model is the suitability for both frontend-backend
likewise backend-backend communication. A service therefore can evolve
from an heavy backend with a lot of network roundtrips to a leaner
backend seamlessly. The browser build-in \textbf{fetch API} which is
essentially a HTTP request can be heavily incorporated into a
browsernative microservice to ensure communication to services in the
backend.

\subsection{2.4. Decentralized
Governance}\label{decentralized-governance}

Microservices are separate entities and decentralization is important to
ensure autonomy. This paper already described fragmented services
bounded to singular business context choreographed by simple
communication protocols developed and evolved by autonomous teams.

This distributed nature empowers teams to create their own technology
stack, tools and services designed in the spirit of language- and
platform independence and share their knowledge with other
parties.{[}4{]} In the recent years many big companies like Facebook,
Google, Netflix and others followed that spirit and published their
ideas and implementations open source. The previously mentioned ReactJS
for example is a product of Facebooks need to ensure a consistent
frontend experience. In fact, many tools and techniques are byproduct of
vital interaction of concrete domain problems and their implementations.

The spirit of freedom can't be applied universally to
\emph{browsernative microservices} as the browser and its underlying DOM
will be the limitation factor to a certain degree. Talking about the
browser, a reader might be tempted to narrowly thinking of the obvious
VIEW layer only - which is not true anymore. In the recent years the
major browser engines grow to to a fully-fledged app deployment platform
offering connectors to build-in databases, multithreading support and
ever-growing JS build-ins like speech synthesis or push notifications.
So-called \textbf{Progressive Web Apps}\footnote{\href{https://developers.google.com/web/progressive-web-apps}{Progressive
  Web Apps}}, a bunch of criteria for building good browser apps, can
achieve a similar look and feel like native apps. And last but not least
services like NativeScript\footnote{\href{https://www.nativescript.org}{NativeScript}}
effectively compiling ``the web'' to native machine code lowering the
boundary between native and browser code even further.

JS is the widely accepted language of the web. Nevertheless, a
microservice engineering team might choose another language for various
reasons. Transpiling languages to JS as target language is a stable
solution nowadays. Languages like TypeScript, ClojureScript or
PureScript compile to JS even exclusively. Once web components hit a
critical mass there will be most likely some library support or foreign
function interface towards ES6 modules (which are mandatory for the new
specifications).

Another more real life decentralization aspect derives from the easiness
of deployment in a safe, sandboxed environment. Web components virtually
ship no overhead or require dedicated build tools. This makes them ideal
candidates for sharing and open source publishing similar to the largest
JS package registry NPM. In the spirit of NPM web components can be
perceived as frontend packages with an HTML interface instead of a JS
signature. Webcomponents.org\footnote{\href{https://webcomponents.org}{Webcomponents.org}}
is a registry for ready-to-use components of every scale and purpose
where for even Google shares a lot of their material design elements.

\subsection{2.5. Decentralized Data
Management}\label{decentralized-data-management}

Data Management in a microservice follows the same modular philosophy
like the service implementation. As mentioned earlier different bounded
contexts make different assumptions of the underlying models. A
\emph{browsernative microservices} takes this idea even further and
expands it to the fragmented world of electronic devices. Decentralized
decisions about conceptual models demand for decentralized data storage
decisions.{[}4{]} Todays web architectures aims to leverage an
increasing amout of processing to the client to avoid time-consuming
roundtrips especially in mobiles networks.\footnote{Latency numbers:
  https://gist.github.com/jboner/2841832} Since network roundtrips are
costly it is a good advice to only query as much data as needed and
cache as much as possible. The build-in LocalStorage or its successor
IndexedDB are mature persistence technologies and libraries like
PouchDB\footnote{\href{https://pouchdb.com/}{PouchDB}} even offer
backend adapters for syncing out of the box.

``Microservices prefer letting each service manage its own
database.''{[}Fowler2014{]} Ben Issa, chief architect of ING Australia
emphasizes this pragmatism on APIs in a conference talk. At ING the
frontend demands tailor the backend APIs, APIs may be produced
automatically and not even Issa knows how many APIs exists.{[}8{]} They
are using a pattern called \textbf{backend for frontends} empowering the
team working to craft their UI and backend in a one-to-one
relationship.{[}3, p. 72{]}

To see this pattern in the field a reader might have a look at Facebooks
GraphQL\footnote{\href{http://graphql.org/}{GraphQL}}. GraphQL is a
query language for the frontend. The backend solely replies on the
frontend needs. Another well documented example in the field is
Cognitects Datomic\footnote{\href{http://www.datomic.com/}{Datomic}},
where parts of the database will be reflected to the client. A so-called
Transactor ensures ACID compliance.

The simplified microservice example later in this paper assumes a
generic build-in API accompanied by build-in frontend components.
Instead of gluing frontend and backend together on runtime the
microservice is designed holistically containing both front- and
backends. For the sake of simplicity data management won't be explored
into depth throughout this paper.

\subsection{2.6. Infrastructure
Automation}\label{infrastructure-automation}

Microservices tend to increase complexity as this model adds a sheer
number of moving parts to the system whereas requires proper
orchestration.{[}3, p. 246{]} Arguably every more sophisticated web
developer already came across build tools like Webpack or infrastructure
automation tools like Gulp. Testing and deploying web components
shouldn't be an obstacle in development.

In the global nature of web development the development couldn't
completely decoupled from the production environment. This circumstance
left developers switching back and forth between files developing tricky
opinionated (and more often biased) ways to glue related parts together.
Bret Victor, UI designer at Apple defined the importance of an
\textbf{immediate feedback principle} for developing user
interfaces.\footnote{\href{https://vimeo.com/36579366}{Bret Victor -
  Inventing on Principle}} In his talk he emphasizes the importance of
an immediate connection between the creator of a product and product
itself. Any change must results in an immediate visible feedback. Web
components catch up with this principle as they allow isolated
development within a single file containing all bits and pieces of the
web component. Every major browser devtool offers a direct file
manipulation functionality so development can be even in place.

When it comes to standardized deployment guidelines previously mentioned
Ben Issa, described the ING standard workflow. Every component comes in
its own \textbf{git repo} containing:

\begin{itemize}
\tightlist
\item
  Internationalization conformity (i18n)
\item
  Accessibility conformity (a11y)
\item
  Tests for the component
\item
  Demos of the component
\item
  Blueprints to mock the one to one APIs
\item
  Docs
\end{itemize}

Even though this example is an opinionated perception it gives a sense
of a mature component build for the web. This example should made clear
that all parts of the component put together in one place. Every
check-in is handled as release candidate and can be independently tested
and deployed by a fully automated machinery.{[}8{]} Due to an exhaustive
amount of testing and deployment tools for JS an automated
infrastructure shouldn't be a problem.

\subsection{2.7. Design for failure}\label{design-for-failure}

In theory a microservice is designed with focus on monitoring for both
the architectural elements and business relevant metrics.{[}4{]} Due to
the modular structure weak points can occur in the orchestration of the
services. A microservice should track down every communication flow and
provide defaults and meaningful error messages where communication
stuck. Testing every single component with predefined synthetic events
ensures functionality. Nevertheless, browser support may vary and legacy
browsers remain a general problem for enhancing websites with new
technologies and therefore demand further configuration.

Combinment of the resources in the browser always demanded for
optimization to avoid unexpected side-effects like \emph{flash of
unstyled content}. Googles Polymer propagates the a general-purpose
pattern called \textbf{PRLP}\footnote{\href{https://www.polymer-project.org/1.0/toolbox/server}{PRLP
  pattern}}:

\begin{itemize}
\tightlist
\item
  Push critical resources for the initial route
\item
  Render initial route
\item
  Pre-cache remaining routes
\item
  Lazy-load and create remaining routes on demand
\end{itemize}

Following this pattern a critical resource can evaluate browser maturity
beforehand and switch to a \textbf{polyfill} or another fallback
solution instead of the latest browser optimized version. After the
initial paint, critical resources like top-level microservices or other
app logic can be loaded and registered.

Regarding the evolution of the web, the ``next billion'' internet users
most likely using Android, have decent specs mobile phones, use an
evergreen browser but won't have a reliable internet connection.{[}9{]}
While \textbf{Progressive Enhancement} was once related to build
websites both for browsers with and without JS support the demands
changed tends towards an \textbf{offline first} principle avoiding
network connectivity failures.{[}9{]} A \emph{browsernative
microservice} therefore not only tries to cache data as much as
possible, it should also bring in a lot of program logic as described in
the previous chapters.

\subsection{2.8. Evolutionary Design}\label{evolutionary-design}

Microservices tend to become smaller over time. An evolutionary design
approach puts emphasizes on decomposition and scrapping the service.
``The key property of a component is the notion of independent
replacement and upgradeability.''{[}4{]} Therefore we can safely change
and chop services. Lazy components of the system which won't change
often should be separated from parts undergoing a lot of churn.{[}4{]}
And services which change for the same reason might be moved together or
even could be merged.

Pursuing flexibility for web development is a selling point as
innovation cycles for browser development is fast paced and technologies
can change quickly. Frontend related hardware, software and
methodologies innovate rapidly over time.

\emph{Browsernative microservices} should be perceived as complementary
technology in contrast to full-service frameworks like Angular. Being
native technology pursues a strong interopt approach with existing
systems. Andrew Rota for example came up with a pattern using small,
encapsulated and stateless web components as leaves in the tree of React
components instead of native HTML elements.{[}10{]} Even React can
eventually profit from the expressiveness of custom components using a
custom \texttt{\textless{}meaningful-button\textgreater{}} over a native
\texttt{\textless{}button\textgreater{}}. Most likely there will be
always some cutting edge framework promising advantages over native
code. Whatever new framework will be on the rise within the next years
native components can eliminate future uncertainty and allowing rapid
reassembling towards new architectures.

\section{3. W3C specifications}\label{w3c-specifications}

For building a native microservice running on the ``bare-metal'' browser
engine requires a bunch of new specifications and assumptions. Most
importantly the quasi specification \textbf{Web Components} is needed.
Web Components is not a real standard. It's an amalgam of APIs from
multiple w3c specs which can be used independently. A web developer may
choose one spec and embrace the freedom in architecture which can be
combined with other frameworks/libraries.

Depending on the context, some people argue for only two specs which
essentially make it possible to create a scoped component but not caring
too much on it's distribution{[}11{]}. Some people prefer the three
specs {[}12{]}, but the majority advocating the four specs variant,
which is listed on the quasi-official
\href{http://webcomponents.org}{webcomponents.org} website. For the
purpose of this article, all four specs will be discussed briefly to
provide a rough understanding. It is not meant to cover all bits and
pieces.

\emph{Disclaimer:} This paper introduces many new browser build-ins with
the focus on testability. As the time of writing, many examples can be
tried frictionless in the console of the latest versions of
\textbf{Google Chrome, Opera and Apple Safari}.\footnote{\href{http://jonrimmer.github.io/are-we-componentized-yet/}{Are
  we componentized yet?}} On Mozilla Firefox technologies work behind a
flag and Microsoft Edge implementation is unfortunately far behind. But
Browser implementation changes quickly and soon technology adoption
won't be an issue. Meanwhile new standards can be used through
\textbf{polyfills} even on legacy browsers.

\subsection{\texorpdfstring{3.1. Custom Elements
\href{https://html.spec.whatwg.org/multipage/scripting.html\#custom-elements}{(whatwg)}}{3.1. Custom Elements (whatwg)}}\label{custom-elements-whatwg}

\emph{Custom elements} are the fundamental building blocks for web
components introducing the \textbf{Single Responsibility Principle} to
the browser. In short, they provide a way to create custom HTML tags
subsuming behavior, design and functionality. An obligatory
\textbf{HelloWorld} will give a flavor about the spec:

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{HelloWorld}\NormalTok{.}\AttributeTok{js}
\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
 \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
  \KeywordTok{super}\NormalTok{()}\OperatorTok{;} \CommentTok{// mandatory in constructor}
  \KeywordTok{this}\NormalTok{.}\AttributeTok{onclick} \OperatorTok{=} \NormalTok{e }\OperatorTok{=>} \AttributeTok{alert}\NormalTok{(}\StringTok{"hello"}\NormalTok{)}\OperatorTok{;}
 \OperatorTok{\}}
\OperatorTok{\}}
\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> index.html}
\KeywordTok{<hello-world>}\NormalTok{say hello}\KeywordTok{</hello-world>}
\end{Highlighting}
\end{Shaded}

This example should be almost self-explanatory in functionality.
\emph{Custom elements} come in the fashion of \textbf{ES6
Classes}\footnote{\href{https://developer.mozilla.org/en/docs/Web/JavaScript/Reference/Classes}{JavaScript
  Classes}} in favor of the normal JavaScript prototype-based
inheritance model which was part of an older specification. Every valid
element must extend the base \texttt{HTMLElement} interface
which~``ensures the newly created element inherits the entire DOM API
and any properties/methods that you add to the class become part of the
element's DOM interface.''{[}13{]} Like any other ES6 class any
\emph{custom element} can be specialized further using the typical
inheritance model allowing higher levels of abstraction.

The beauty of \emph{custom elements} comes with the \textbf{bounded
\texttt{this} keyword} which points to the element itself. Instead of
querying and assigning behavior after creation of the node, custom
elements ship their functionality on initialization of the element. The
so called \emph{fat-arrow} (\texttt{=\textgreater{}}) is just a new ES6
syntax feature for an anonymous function declaration.

After declaration the new HTML element needs to be registered in the
global build-in \texttt{customElements} object with an dedicated tag
name acting as key to the element. Mind the dash inside the tag name to
conform the spec. Finally, the new element can be mounted inside the
HTML document.

\subsubsection{3.1.1. Lifecycle methods}\label{lifecycle-methods}

In addition to the constructor which runs procedures on initialization,
the spec defines \textbf{lifecycle callbacks} for controlling elements
behavior towards DOM interaction. Many popular frameworks like React or
Angular rely on similar approaches:

\begin{itemize}
\tightlist
\item
  connectedCallback()\\
  Called upon the time of \textbf{connecting or upgrading the node}
  which means the moment the node is rendered inside the DOM. Typically
  this method is called straight after the constructor if the node is
  inserted directly. For a faster initial render of the page it is
  highly preferable to put many proceedings in this method. Usually this
  method contains setup code such as fetching resources or rendering
  elements according to attributes.{[}13{]}
\item
  disconnectedCallback()\\
  Called upon the time of \textbf{node removal}. Cleanup code like
  removing event listeners or disconnecting web sockets can be put here.
\item
  attributeChangedCallback(attrName, oldVal, newVal)\\
  This method provides an \textbf{onchange handler} for certain elements
  attributes. This method is used to guide elements' transition from an
  old value to a new state. Due to performance issues this callback is
  only triggered for attributes registered in a dedicated array shipped
  with the element.
\item
  adoptedCallback()\\
  Called when moving the node \textbf{between documents}. This method
  comes handy when using HTML imports described later.
\end{itemize}

\subsubsection{3.1.2. Custom attributes}\label{custom-attributes}

As mentioned earlier any custom elements must extend the
\texttt{HTMLElement} interface ensuring base properties and methods used
in throughout all HTML elements like \textbf{id, class, addEventListner
\ldots{}}. Additionally, it is possible to define custom attributes
using the \emph{custom elements'} \textbf{getter / setter interface} to
steer the behavior of the element. Note that the keywords \textbf{get}
and \textbf{set} as well as the previously used \textbf{constructor} are
optional!

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{HelloWorld}\NormalTok{.}\AttributeTok{js}
\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
  \NormalTok{set }\AttributeTok{sayhello}\NormalTok{(val) }\OperatorTok{\{}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello} \OperatorTok{=} \NormalTok{val}\OperatorTok{;}
  \OperatorTok{\}}
  \NormalTok{get }\AttributeTok{sayhello}\NormalTok{() }\OperatorTok{\{}
    \ControlFlowTok{return} \KeywordTok{this}\NormalTok{.}\AttributeTok{_hello}\OperatorTok{;}
  \OperatorTok{\}}
\OperatorTok{\}}
\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
\CommentTok{// Instantiation via JS instead of HTML}
\KeywordTok{var} \NormalTok{el }\OperatorTok{=} \KeywordTok{new} \AttributeTok{HelloWorld}\NormalTok{()}\OperatorTok{;}
\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello} \OperatorTok{=} \StringTok{"earth"}\OperatorTok{;} \CommentTok{// "Call" the setter}
\VariableTok{el}\NormalTok{.}\AttributeTok{sayhello}\OperatorTok{;}\CommentTok{// Yields "earth"}
\end{Highlighting}
\end{Shaded}

Native DOM properties always reflect their values between HTML and
JS.{[}14, Para. 2.6.1{]} Declaring
\texttt{\textless{}hello-world\ id="hello"\textgreater{}} equals to the
JS declaration \texttt{new\ HelloWorld().id\ =\ "hello"}.

This behavior won't work out-of-the-box with methods defined with
setters as they are strictly JS only . Mounting
\texttt{\textless{}hello-world\ sayhello="mars"\textgreater{}} would't
call the \texttt{sayhello} method in the previous setup. Value
reflection can be implemented inside the \emph{custom elements} using
the native methods \texttt{getAttributes} and \texttt{setAttributes}.
Using them exhaustively throughout lifecycles methods the new components
can configured to read and listen to HTML attributes accordingly.

Designing a \emph{custom element} this way creates HTML elements with
named attribute interfaces reaching deep into JS functionality. With
this mental model in mind a web developer can create highly dynamic web
components.

\subsubsection{3.1.3. Customized build-in
elements}\label{customized-build-in-elements}

One aspect didn't mentioned yet is the possibility of extending other
build-in elements by extending other interfaces instead of the
\texttt{HTMLElement} interface. While this functionality is perfectly
spec'd it is strongly rejected by some browser vendors.\footnote{https://github.com/w3c/webcomponents/issues/509}
Most likely the spec will change in future in one or other way on this
issue and therefore customized build-in elements left out of this paper
intentionally.

\subsection{\texorpdfstring{3.2. Shadow DOM
\href{http://w3c.github.io/webcomponents/spec/shadow/}{(w3c)}}{3.2. Shadow DOM (w3c)}}\label{shadow-dom-w3c}

A \emph{shadow DOM} is just an isolated DOM tree living inside an
another DOM tree. The spec refers the hosting tree as \textbf{light DOM
tree} and the attached DOM as \textbf{shadow DOM tree}. Conceptually
\emph{shadow DOM} issues a single important issue for building scalable
software which is namely \textbf{encapsulation}. While custom elements
provide a good way to encapsulate JS behavior \emph{shadow DOM} tends
strongly to the direction of style and event encapsulation.

With an ever increasing complexity of single-page applications the
global nature of the DOM creates a daunting situation for code
organization and leads over times to highly fragmented bits of CSS and
obscured CSS selectors. Of course this situation lowers code clarity and
reusability dramatically. The only solution which won't break with the
existing global paradigm of the DOM effectively is to allow separate
pieces of encapsulated code sit on top of the global DOM - introducing
the shadowed DOM approach.

Enhancing the previous example the new encapsulated \texttt{HelloWorld}
would like this:

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{HelloWorld}\NormalTok{.}\AttributeTok{js}
\KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
 \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
   \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
   \VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{innerHTML} \OperatorTok{=} \StringTok{'<p>hello</p>'}\OperatorTok{;}
 \OperatorTok{\}}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

The new global method \texttt{attachShadow} adds a new document root to
the \texttt{HelloWorld} which has the same properties as a normal, light
document object. Note that the \texttt{shadowRoot} object is marked as
\textbf{open} which ensures that some events can bubble out and outside
JS can reach in the new root.

Filling the \emph{shadow DOM} with an \texttt{innerHTML} string is
rather impractical. To fill a \emph{shadow DOM} with life usually it
invites light DOM child nodes nested under the hosting node using a
technique called \texttt{slots}.

\subsubsection{3.2.1. Slots}\label{slots}

Contradicting to the simplified \texttt{HelloWorld} example, a
\emph{shadow DOM} shouldn't contain dynamic content. Changing or
interacting with the paragraph node from the example would require
nested JS calls querying the hosting node, entering the \emph{shadow
DOM} and applying a function. Imported JS behavior from third-party
libraries in the \emph{light DOM} can't be used inside the \emph{shadow
DOM}, too.

That's why the shadowed documents should be more perceived as
\textbf{static documents} filled and managed solely by the render
engine. \texttt{Slots} are target areas for \emph{light DOM} nodes used
to mark the endpoints in question.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> index.html}
\KeywordTok{<HelloWorld-with-ShadowDOM>}
  \CommentTok{<!--}
\CommentTok{    All child nodes will be moved inside the}
\CommentTok{    shadowRoot if shadowRoot.innerHTML = '<slot></slot>'}
\CommentTok{    -->}
  \KeywordTok{<p>}\NormalTok{hello I will be scoped}\KeywordTok{</p>}
\KeywordTok{</HelloWorld-with-ShadowDOM>}
\end{Highlighting}
\end{Shaded}

Technically, the \emph{light DOM} nodes are not moved inside the
\emph{shadow DOM}. Their just rendered in place. It's an subtle but
important difference towards handling a node. All JS behavior and CSS
styles applied in the \emph{light DOM} will be valid in the \emph{shadow
DOM}. The render engine literally taking the nodes and drop them inside
the \texttt{slot} tag. This procedure is commonly referred as
\textbf{flattening} of the DOM trees.

It's possible to add semantics to the \emph{shadow DOM} in naming the
slots which frees the \emph{light DOM} from the responsiblity to deliver
nodes in a correct top-to-bottom order. Combining \emph{shadow DOM} with
HTML templates provides the web developer with a flexible \textbf{HTML
template engine}.

\begin{longtable}[]{@{}ll@{}}
\toprule
FROM: light DOM & TO: shadow DOM\tabularnewline
\midrule
\endhead
\texttt{\textless{}p\ slot="hello"\textgreater{}Named\textless{}/p\textgreater{}}
&
\texttt{\textless{}slot\ name="hello"\textgreater{}hello\ only\textless{}/slot\textgreater{}}\tabularnewline
\texttt{\textless{}p\textgreater{}Unnamed\textless{}/p\textgreater{}} &
\texttt{\textless{}slot\textgreater{}Unnamed\ nodes\textless{}/slot\textgreater{}}\tabularnewline
\bottomrule
\end{longtable}

Writing a little documentation inside the
\texttt{\textless{}slot\textgreater{}} tag is considered as a good
practice as it provides the developer with visual clues what nodes must
be delivered. This functionality makes a \emph{shadow DOM} pretty much
self-explanatory. Inside a default slot tag the render engine expands
all \emph{light DOM} children without a named \texttt{slot} attribution.

\subsubsection{3.2.2. Styling}\label{styling}

As mentioned in the last section, there is a distinct difference about
the nature of nodes. Nodes declared and rendered exclusively in the
\emph{shadow DOM} are not affected by any styling from outside. Nodes
which are distributed will be styled in the \emph{light DOM} and can be
additionally painted in the \emph{shadow DOM}.

Note that styles from the outside have an higher specify than styles
assigned after distribution. Therefore it is generally a good advice to
minimize the global stylings to some base styling for uniformity of the
web site while leaving the specific stylings to the component. It is
possible to \textbf{reset all styles} inside the \emph{light DOM} before
distributing nodes using the \textbf{\texttt{all:\ initial}} reset. To
ensure a consistent look between different shadow roots this technique
should be used carefully.

Regarding the importance style encapsulation, a couple of \textbf{new
CSS rules} emerged that are exclusively targeting the \emph{shadow DOM}.
The table below outlines styling possibilities for the use INSIDE the
\emph{shadow DOM}:

\begin{itemize}
\tightlist
\item
  ::slotted(selector)\\
  Applies to distributed nodes and repaints them after distribution.
  Slotted won't override outsides styles but can complement them
  previously unset style rules.
\item
  :host\\
  The host property will add styles or change inherited ones inside
  shadow DOM. Aforementioned style resets can be placed here.
\item
  :host(condition)\\
  Like the previous rule this selector will style the shadow DOM but
  this time based on attributes/conditions assigned to the hosting node.
\item
  :host-context(condition)\\
  Like the previous rule this selector will style the shadow DOM but
  will look after context set at the host node or even at the host
  ancestor.
\end{itemize}

Using the \textbf{functional selectors} \texttt{host()} or
\texttt{host-context()} allows the creation of context-aware custom
elements. A possible usecase would be ``theming'' a component.

\subsubsection{3.2.3. Behavior}\label{behavior}

As mentioned earlier any logic applied to \emph{light DOM} nodes stays
with the node even after redistribution. For the sake of separation of
concerns the business logic should be part of the \emph{custom element}
(the \emph{light DOM}) and not the part of the \emph{shadow DOM}. On the
other hand there are numerous scenarios where JS is used for styling or
animation of an element. In this case it might be more straightforward
to \textbf{apply JS inside the \emph{shadow DOM}} to avoid mixing with
logic handlers with styling.

It seems that a \emph{light DOM} node is in the \emph{shadow DOM}
context after distribution. Visually this may be true but logically the
node stays in the normal document. To reach a distributed node from the
\emph{shadow DOM} to apply some JS behavior for styling it needs the
extra way over the slot node. Calling \texttt{assinedNodes()} on the
hosting slot element returns a linkage to the distributed node which can
be accessed and manipulated like in the \emph{light DOM} context.

Wrapping up this section \emph{shadow DOM} provides a non-hacky way to
create uniform looking custom elements and even enhance styling
possibilities without adding overhead. For small components with just a
little styling a \emph{shadow DOM} might be over engineered. Eventually
it all depends on the question of ``how hard is it to implement it
without shadow DOM'' - which can't be answered universally. For a more
in-depth guide, Google Engineer Eric Bidelman wroten a great primer on
\emph{shadow DOM}{[}15{]}.

There is still a missing link between \emph{light DOM} and \emph{shadow
DOM}. The observant reader may have already noticed the weak point in
the HelloWorld example: how to ``vitalize'' the \emph{shadow DOM} with
slot or other element structuring instead of \texttt{innerHTML} strings.
While strings works perfectly fine in this simple case a string of
markup is rather cumbersome and error-prone and doesn't scale well. When
putting quotes inside other quotes things break quickly. Strings make
development harder because code editor features like indentation or
syntax highlighting won't be supported. The HTML templates is set to
fill this gap.

\subsection{\texorpdfstring{3.3. HTML Templates
\href{https://html.spec.whatwg.org/multipage/scripting.html\#the-template-element}{(whatwg)}}{3.3. HTML Templates (whatwg)}}\label{html-templates-whatwg}

Among all other new specifications \emph{HTML templates} are the most
mature and adopted standard in the browser environment. All major
browsers support it since years.

One core concept in templates is browser performance. Elements inside a
\texttt{template} tag will be parsed on runtime - but not constructed
and rendered into the content tree. They're remaining plain HTML markup
sitting in the document until the time of activation.

Activation usually takes four steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Querying the template node in question}\\
  const node = document.querySelector(`template');
\item
  \textbf{Parsing the content and preparing the templates' content}\\
  const content = node.content;\\
  Returns a DocumentFragment object.
\item
  \textbf{Optional: Cloning the fragment for multiple use}\\
  const clone = content.cloneNode(``deep'');
\item
  \textbf{Appending the clone/original to destination}\\
  document.body.appendChild(clone);
\end{enumerate}

As easy and minimal \emph{HTML templates} are they're missing out a
crucial feature other template implementations used to have. As
templates are basically just containers for HTML markup there is no
idiomatic way to define \textbf{placeholders} for dynamic content.
Templates could be mock up this way using JS for altering the content
but the much cleaner way leverages the previously described slot
technique from shadow DOM.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> hello-world-component.html}
\KeywordTok{<hello-world>}
  \KeywordTok{<template}\OtherTok{ id=}\StringTok{"hello"}\KeywordTok{>}
  \CommentTok{<!-- styling -->}
    \KeywordTok{<style>}
      \FloatTok{#hellowrap} \KeywordTok{\{}
        \KeywordTok{font-weight:} \DataTypeTok{bold}\KeywordTok{;}
        \KeywordTok{color:} \NormalTok{orange}\KeywordTok{;}
      \KeywordTok{\}}
    \KeywordTok{</style>}

    \CommentTok{<!-- structure -->}
    \KeywordTok{<div}\OtherTok{ id=}\StringTok{"hellowrap"}\KeywordTok{>}
      \KeywordTok{<slot}\OtherTok{ name=}\StringTok{"placeholder"}\KeywordTok{>}
        \NormalTok{Named placeholder}
      \KeywordTok{</slot>}
    \KeywordTok{</div>}
  \KeywordTok{</template>}

  \CommentTok{<!-- light DOM / typical inserted at index.html -->}
  \KeywordTok{<p}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
    \NormalTok{Hello World Web Component}
  \KeywordTok{</p>}
\KeywordTok{</hello-world>}

\KeywordTok{<script>}
  \KeywordTok{class} \NormalTok{HelloWorld }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
    \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
      \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
      \KeywordTok{this}\NormalTok{.}\AttributeTok{attachShadow}\NormalTok{(}\OperatorTok{\{}\DataTypeTok{mode}\OperatorTok{:} \StringTok{'open'}\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
      \KeywordTok{const} \NormalTok{temp }\OperatorTok{=} \KeywordTok{this}\NormalTok{.}\AttributeTok{querySelector}\NormalTok{(}\StringTok{'template#hello'}\NormalTok{)}\OperatorTok{;}
      \KeywordTok{this}\NormalTok{.}\VariableTok{shadowRoot}\NormalTok{.}\AttributeTok{appendChild}\NormalTok{(}\VariableTok{temp}\NormalTok{.}\AttributeTok{content}\NormalTok{)}\OperatorTok{;}
    \OperatorTok{\}}
  \OperatorTok{\}}
  \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} \NormalTok{HelloWorld)}\OperatorTok{;}
\OperatorTok{<}\SpecialStringTok{/script>}
\end{Highlighting}
\end{Shaded}

The updated \texttt{HelloWorld} component looks already pretty mature.
It combines all the previous mentioned standards into one HTML file.
Custom elements serves the logic, shadow DOM scopes the styles and
\emph{HTML Templates} efficiently glues light DOM and shadow DOM
together. This separation of concerns comes with a surplus in
flexibility. In a real world scenario \texttt{HelloWorld} could
reference multiple \emph{HTML Templates} and switch them around without
any fuss. Even further a developer might split up templates into named
\textbf{STYLE} templates and \textbf{CONTENT} templates to increase
reusability even further.

The last standard in the row of four is not concerned with the internals
implementation of a web component. HTML Imports serves the need for an
efficient distribution mechanism of components and other HTML resources.

\subsection{\texorpdfstring{3.4. HTML Imports
\href{https://www.w3.org/TR/html-imports/}{(w3c)}}{3.4. HTML Imports (w3c)}}\label{html-imports-w3c}

Importing the \texttt{HelloWorld} component is a one-liner:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> index.html}
\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"Hello.html"}\OtherTok{ async}\KeywordTok{>}

\KeywordTok{<hello-world>}
  \KeywordTok{<p}\OtherTok{ slot=}\StringTok{"placeholder"}\KeywordTok{>}
    \NormalTok{Hello World Web Component}
  \KeywordTok{</p>}
\KeywordTok{</hello-world>}
\end{Highlighting}
\end{Shaded}

The \texttt{async} flag is optional but recommended like any other
fetching event. Once the imported HTML document comes into scope
activation follows a very similar process compared to the aforementioned
HTML templates:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Querying the link node}
\item
  \textbf{Parsing the content and preparing the render}\\
  const content = linknode.import; -\textgreater{} Unlike the \emph{HTML
  template} a complete document object is constructed.
\item
  \textbf{Optional: Cloning some nodes for multiple use}
\item
  \textbf{Appending the clone/original to destination}
\end{enumerate}

Again this is the imperative way to handle a generic \emph{HTML Import}.
In the declarative world of web components a component is
\textbf{parsed, auto-activated and anchored} solely by its' tag name
\texttt{\textless{}hello-world\textgreater{}\textless{}/hello-world\textgreater{}}
on purpose. The very own lifecycle method \textbf{adoptedCallback()} in
custom elements shows the strong interconnection between those
standards.

\emph{HTML imports} can import everything which is wrappable with HTML
markup. Stylesheets, scripts, documents, media files and even further
imports statements can form a semantic \emph{HTML import} statement. The
far reaching possibilities of a single standard has its downsides.
According to Mozilla \emph{HTML imports} are not fully compatible with
the dependency model of the new ES6 modules.\footnote{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}
Under the current situation Mozilla and Apple and Microsoft won't
implement \emph{HTML Imports} soon if any. Currently only Googles blink
web engine supports \emph{HTML Imports} as they are the driving force
behind the web components specs in general. Despite the discrepancies
among browser vendors \emph{HTML Imports} are part of this paper as no
other native standard can ship bundled HTML, CSS and JS which is a core
concept behind web components.

\subsection{\texorpdfstring{3.5. Custom Events
\href{https://dom.spec.whatwg.org/\#interface-customevent}{(whatwg)}}{3.5. Custom Events (whatwg)}}\label{custom-events-whatwg}

Events are first-class citizens in the browser. After all they provide a
neat communication channel for dynamic interactions. \emph{Custom
Events} are part of the DOM since years but with the rise of web
components they will most likely become an indispensable building block
of web components.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> hello-world-component.html}

\KeywordTok{<hello-world>}
 \KeywordTok{<button>}\NormalTok{Launch CustomEvent}\KeywordTok{</button>}
\KeywordTok{</hello-world>}

\KeywordTok{<script>}
 \VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'hello-world'}\OperatorTok{,} 
 \KeywordTok{class} \KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
    \KeywordTok{super}\NormalTok{()}\OperatorTok{;}

   \CommentTok{// Craft a helloevent}
    \KeywordTok{const} \NormalTok{helloevent }\OperatorTok{=} \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'helloevent'}\OperatorTok{,} \OperatorTok{\{}
    \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}
    \DataTypeTok{detail}\OperatorTok{:} \StringTok{'Contains scalar or object'}
   \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}

   \CommentTok{// Launch helloevent if child button clicks}
   \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'click'}\OperatorTok{,} \NormalTok{click }\OperatorTok{=>} \OperatorTok{\{}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(helloevent)}
    \VariableTok{click}\NormalTok{.}\AttributeTok{stopPropagation}\NormalTok{()}\OperatorTok{;}
   \OperatorTok{\}}\NormalTok{)}
  \OperatorTok{\}}
 \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
\OperatorTok{<}\SpecialStringTok{/script>}
\end{Highlighting}
\end{Shaded}

Naming events after the emitting tag makes the API almost
self-explanatory. The \texttt{detail} property can be loaded with
scalars as well as objects. Subsequent parent nodes may catch the custom
event with a clear understanding about the source node.

Chaining and aggregating events from child nodes can be frequently used
within web components. One use case could be creating a custom button
like in the previous example. As mentioned earlier in the Custom
Elements section at 3.1.3. , the pattern of \textbf{extending native
elements} should be somewhat dismissed as other browser vendors may
never implement it. A common workaround to eventually create a own
version a button could be made with a thin wrapper around the native
button and chaining a \emph{Custom Event} after the click event.

Another common web components use case can be a \textbf{middleware}
subscribing to certain child events and acting upon them. This involves
catching events, buffering, destructuring and creation of an own event
towards the document root. By design, events only bubble upstream
towards parent nodes. For handing down information towards the child
nodes we need to query the child node directly.

\subsection{\texorpdfstring{3.6. Web Worker
\href{https://html.spec.whatwg.org/multipage/workers.html}{(whatwg)}}{3.6. Web Worker (whatwg)}}\label{web-worker-whatwg}

Like \emph{Custom Events}, \emph{Web Workers} had been around for a long
time and therefore enjoy full support among major browsers. They emerged
at around 2009 when discussions about browser performance was still in
the early days but addressing fundamental performance bottleneck of the
JS language.

JS runs in a single-threaded language environment. Every script in the
browser environment, from handling UI events to query and process larget
amounts of data or manipulating the DOM runs on a single thread{[}16{]}.
Putting a lot of work to the single main thread can slow down the web
service significantly. From time to time scripts can block or fail for
whatever reason which leads to a frozen or crashed UI. A worker can
overcome the bottleneck of the single-threaded nature with spawning new
\textbf{background threads} which allows the UI to stay responsive even
when computation-heavy tasks needs to perform. Furthermore, a worker
thread adds a performance advantage embracing the multi core CPU
architecture most devices running on today. To grasp the full potential
of workers a reader might dive deeper into the Angular 2 architecture,
where most of the application layer is abstracted from the main
rendering thread into worker threads.\footnote{\href{https://docs.google.com/document/d/1M9FmT05Q6qpsjgvH1XvCm840yn2eWEg0PMskSQz7k4E}{Angular
  2 Rendering Architecture}}

A \emph{Web Worker} spawns a new background thread where scripts can run
concurrent to the main thread. Usually a worker is loaded from dedicated
file to embrace separation.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{const} \NormalTok{worker }\OperatorTok{=} \KeywordTok{new} \AttributeTok{Worker}\NormalTok{(}\StringTok{'worker.js'}\NormalTok{)}\OperatorTok{;}
\end{Highlighting}
\end{Shaded}

After initialization a worker communicates over a simple \textbf{message
based interface} with the main thread.

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{main}\NormalTok{.}\AttributeTok{js}
\CommentTok{// Send to worker}
\VariableTok{worker}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{'Hello World'}\NormalTok{)}\OperatorTok{;}
\CommentTok{// Receive msg from worker}
\VariableTok{worker}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
 \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\StringTok{'Worker said: '}\OperatorTok{,} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{>} \VariableTok{worker}\NormalTok{.}\AttributeTok{js}
\CommentTok{// Receive msg and echo back}
\KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>}
 \KeywordTok{this}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(}\StringTok{"Echo "} \OperatorTok{+} \VariableTok{e}\NormalTok{.}\AttributeTok{data}\NormalTok{))}\OperatorTok{;}
\end{Highlighting}
\end{Shaded}

\section{4. Building a browsernative
microservice}\label{building-a-browsernative-microservice}

After getting confidence in microservice principles and technical
background the paper should briefly join them to form a
\emph{browsernative microservices}. Needless to say the following
example is overall simplified towards illustrating the connection
between browsernative technologies and microservice patterns. Googles
library Polymer is a good place for learning about web components in
depth and make use of their toolbox. One of their most famous proof of
concept is the so-called
\href{https://shop.polymer-project.org/}{Polymer Shop} which is a
fully-fledged online shop nested within a single root element
\texttt{\textless{}shop-app\textgreater{}}. This app made of several
main views and many more invisible wrapper elements for routing, service
worker caching, theming, etc. The whole shop runs as a single
application fetching and updating remote resources and switching views.
Let's assume we work in a sales engineering team of the Polymer Shop and
need to rebuild the checkout microservice.

The current checkout can be found at
https://shop.polymer-project.org/checkout. At the time of writing the
checkout is a single, 671 lines of code long Polymer component including
all required fields for sign in, shipping, billing and summarizing the
order. In the spirit of microservices we will split up the microservice
into fine grained components. The shopping cart data is pulled out of a
local storage JSON entity set up previously by another custom-element.

By breaking down the service the \textbf{team defined the business
boundaries} within the checkout process and ended up with following
granular service blocks:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Sign in or Sign up
\item
  Shipping details
\item
  Payment details
\item
  Review and place order
\end{enumerate}

Translated into a raw \textbf{Custom Element} HTML structure, the
top-level microservice might look like the following snippet:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{>shop-checkout.html}
\KeywordTok{<shop-checkout>}
  \KeywordTok{<sign-in></sign-in>}
  \KeywordTok{<shipping-details></shipping-details>}
  \KeywordTok{<payment-details></payment-details>}
  \KeywordTok{<place-order></place-order>}
\KeywordTok{</shop-checkout>}
\end{Highlighting}
\end{Shaded}

Yet already we see the simplicity arouse from web components as they
persue a clear markup. Each of the child components may act
independently over other child nodes utilizing the \textbf{loose
coupling principle}. Each child ships all the HTML, CSS and JS code
needed to fulfil its work following the \textbf{high cohesion
principle}. Each component may contain different views to accommodate
different \textbf{bounded contexts resulting from different devices}.
And last but not least, all of them communicate over an
\textbf{unobtrusive message bus} via the service root component
\texttt{\textless{}shop-checkout\textgreater{}}.

Before diving deeper into implementation, its worth to clarify an
\textbf{architectural pattern} behind components. Any reader of the
paper came across ReactJS / Redux, the concept of components may look
familiar. Dan Abramov, the creator of Redux, once defined a simple
dichotomous pattern for creating UI components.

Firstly, Abramov defined a pattern around \textbf{presentational
components} only related with the concern about \emph{how things look}.
This component literally doesn't know anything about the service in
question which makes the component highly flexible and reusable. They
are controlled solely from the outside, receiving data and dispatching
unbiased events on user interaction.{[}17{]} Most probably every
presentational component embodies more HTML/CSS markup and less JS code.
It should encapsulate its styles from bleeding out and protect its
styles being overwritten. Furthermore, it may contain several templates
to change it's look on different demands.

Secondly, Abramov described components he refers as \textbf{containers}.
A container component is concerned with \emph{how things work}.{[}17{]}
Containers acts as invisible wrappers around presentational components
acting in the sense of UNIX filters. Their job is to fetch data from
child nodes, aggregating events, interacting with the model and push
state back to the presentational components. Consequently they might
contain more JS and less if any HTML markup. We probably don't need to
utilize Shadow DOM as no styles are involved.

Last but not least, the pattern can expanded for illustrational purposes
to \textbf{native components} which is every build-in HTMLElement like
the HTMLButtonElement. Native components are mostly deep nested elements
providing the actual functionality in the browser UI. They are solely
controllable and styleable from the outside and are therefore wrapped in
presentational components and/or containers.

Lets start the service description top-down beginning with the
\textbf{root container} managing the overall service.

\subsection{Root container}\label{root-container}

The root container \texttt{\textless{}shop-checkout\textgreater{}} is
basically just an encapsulation layer in terms of service
functionalities. Encapsulation of CSS won't be necessary at this point
as no styling is involved. A simplified root container for the checkout
might look like the following code snippet.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{> shop-checkout.html}
\CommentTok{<!-- IMPORTS -->}
\KeywordTok{<link}\OtherTok{ rel=}\StringTok{"import"}\OtherTok{ href=}\StringTok{"sign-in.html"}\OtherTok{ async}\KeywordTok{>}
\NormalTok{...}
\CommentTok{<!-- VIEW -->}
\KeywordTok{<shop-checkout>}
  \KeywordTok{<sign-in></sign-in>}
  \NormalTok{...}
\KeywordTok{</shop-checkout>}

\CommentTok{<!-- CONTROLLER -->}
\KeywordTok{<script>}
\KeywordTok{class} \NormalTok{ShopCheckout }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
  \AttributeTok{constructor}\NormalTok{() }\OperatorTok{\{}
    \KeywordTok{super}\NormalTok{()}\OperatorTok{;}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{MODEL} \OperatorTok{=} \KeywordTok{new} \AttributeTok{Worker}\NormalTok{(}\StringTok{'checkout-model.js'}\NormalTok{)}\OperatorTok{;}
  \OperatorTok{\}}

  \AttributeTok{connectedCallback}\NormalTok{() }\OperatorTok{\{}
    \CommentTok{// listen to msg from MODEL}
    \KeywordTok{this}\NormalTok{.}\VariableTok{MODEL}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'message'}\OperatorTok{,} \NormalTok{[Msg Handler])}
    \CommentTok{// listen to child nodes}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{addEventListener}\NormalTok{(}\StringTok{'checkout'}\OperatorTok{,} \NormalTok{e }\OperatorTok{=>} \OperatorTok{\{}
      \CommentTok{// delegate event to handler}
      \KeywordTok{this}\NormalTok{.}\AttributeTok{_sendToMODEL} \OperatorTok{=} \NormalTok{e}\OperatorTok{;} 
    \OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
  \OperatorTok{\}}

  \NormalTok{set }\AttributeTok{_sendToMODEL}\NormalTok{(e) }\OperatorTok{\{}
    \KeywordTok{const} \NormalTok{id }\OperatorTok{=} \VariableTok{e}\NormalTok{.}\VariableTok{target}\NormalTok{.}\AttributeTok{id}\OperatorTok{,} 
          \NormalTok{name }\OperatorTok{=} \VariableTok{e}\NormalTok{.}\VariableTok{target}\NormalTok{.}\AttributeTok{localName}\OperatorTok{,}
          \NormalTok{load }\OperatorTok{=} \VariableTok{e}\NormalTok{.}\AttributeTok{detail}\OperatorTok{,}
          \NormalTok{letter }\OperatorTok{=} \VariableTok{Object}\NormalTok{.}\AttributeTok{assign}\NormalTok{(}\OperatorTok{\{\},} \OperatorTok{\{} \NormalTok{id}\OperatorTok{,} \NormalTok{name}\OperatorTok{,} \NormalTok{load }\OperatorTok{\}}\NormalTok{)}\OperatorTok{;}
    \CommentTok{// send to MODEL}
    \KeywordTok{this}\NormalTok{.}\VariableTok{MODEL}\NormalTok{.}\AttributeTok{postMessage}\NormalTok{(letter)}\OperatorTok{;} 
  \OperatorTok{\}}
  \NormalTok{...}
\OperatorTok{\}}    
\VariableTok{customElements}\NormalTok{.}\AttributeTok{define}\NormalTok{(}\StringTok{'shop-checkout'}\OperatorTok{,} \NormalTok{shopCheckout)}\OperatorTok{;}
\OperatorTok{<}\SpecialStringTok{/script>}
\end{Highlighting}
\end{Shaded}

The purpose of this simplified code snippet is to outline the
\textbf{MVC threefold} in the service root. The MODEL is pushed into the
worker thread, the CONTROLLER in the custom element and the VIEW in the
HTML structuring. On runtime the root container acts like a
\textbf{message dispatcher} implementing the microservice principle of
smart endpoints and dumb pipes.

In order to interact with the MODEL every child node must implement the
\textbf{dedicated \texttt{checkout}} custom event (which will be
explained further on in the next section). Summarizing the
unidirectional communication from the service root perspective looks
like following plot:

\begin{verbatim}
VIEW thread                 |               MODEL thread
                            |
+----------+                |               +----------+
| Checkout | ----------Event Msg--------->  |  Msg     |
|  Event   |                |               | Handler  |
+----------+                |               +----------+
                            |                   |
                            |                   |
                            |               +----------+
                            |               | Generate |
                            |               |  Effect  |
                            |               +----------+
                            |                   |
                            |                   |
+----------+                |               +----------+
|  Msg     |                |               | Postmsg  |
| Handler  | <--------Action Msg----------  |  Action  |
+----------+                |               +----------+    
                            |
\end{verbatim}

Effects are yielded by the asynchronous operation of messages and create
actions returned to sender. Effects can be created by external messages,
like subscription to an WebSocket, too. Effects may be created with
additional resources from the server or syncing with local storage like
in the polymer-shop.

While the msg handlers are basic ``dumb'' switch statements the
\textbf{smartness} solely arouse from intelligent controllers processing
the message. This communication model offers lots of possibilities for
\textbf{evolutionary design} as child nodes can be loosely dropped.
Every child node can be expanded into another full microservice or split
up into separate nodes without notice of the service root. Communication
might \textbf{fail graceful} providing a default console log in case
switches aren't defined yet.

\subsection{Container components}\label{container-components}

The second layer of the checkout microservice like
\texttt{\textless{}sign-in\textgreater{}} or
\texttt{\textless{}shipping-details\textgreater{}} still contain a lot
of logic and no or less styling. Their job is to control their
underlying presentational leaf components, aggregating events and to
define a \textbf{set of actions towards the model} . Every container
mounted directly under the service root may have a dedicated area inside
the worker thread where it's fulfil his duties. For example, a typical
\texttt{\textless{}sign-in\textgreater{}} contains merely two fields
username and password and a submit button. The component listens for the
submit action, aggregating the credentials and might add some semantics
to it like \emph{action: `SIGNIN SUBMITTED'}. Fields and action message
will be dispatched towards the model for further processing like
initiating a authorization process.

Every container component is eligible to aggregate subordinate events
from their children, buffer them and interact with the model trough via
a \textbf{unified message system}. A base class, from which
\texttt{\textless{}sign-in\textgreater{}} or
\texttt{\textless{}shipping-details\textgreater{}} can be extended might
look like this:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{class} \NormalTok{SimpleContainer }\KeywordTok{extends} \NormalTok{HTMLElement }\OperatorTok{\{}
  \CommentTok{// constructor, static methods ...}
  \NormalTok{set }\AttributeTok{_dispatch}\NormalTok{(msg) }\OperatorTok{\{}
    \KeywordTok{this}\NormalTok{.}\AttributeTok{dispatchEvent}\NormalTok{(}
      \KeywordTok{new} \AttributeTok{CustomEvent}\NormalTok{(}\StringTok{'checkout'}\OperatorTok{,} \OperatorTok{\{}
      \DataTypeTok{bubbles}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}
      \DataTypeTok{detail}\OperatorTok{:} \NormalTok{msg}
      \OperatorTok{\}}\NormalTok{)}
    \NormalTok{)}\OperatorTok{;}
  \OperatorTok{\}}
  \NormalTok{set }\AttributeTok{_receive}\NormalTok{(msg) }\OperatorTok{\{}
  \CommentTok{// switch to methods}
    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(}\KeywordTok{this}\NormalTok{.}\AttributeTok{localName} \OperatorTok{+} \StringTok{"received msg"}\NormalTok{)}\OperatorTok{;}
    \VariableTok{console}\NormalTok{.}\AttributeTok{log}\NormalTok{(msg)}\OperatorTok{;}
  \OperatorTok{\}}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

Extending the \texttt{SimpleContainer} will equip every container node
with the unified message interface. Calling
\texttt{this.\_dispatch(msg)} within the container will trigger an event
bubbling upstream. The service root will implement a simple event
listener for \texttt{checkout} events. After receiving an answer from
the MODEL a service root can query the container child node in question
and push forward the answer over the \texttt{\_receive} property.

There might be even more middleware containers pulled in between
presentational leafs and the service root to fulfil some extra work
either like filters on the event or without caring about checkout events
altogether. Changing or enhancing any functionality requires only the
controllers in the container in question and the endpoint section at the
model. \textbf{Infrastructure automation} might be achieved by
dynamically evaluating the mounted containers beneath the service root
and modeling the ``backend'' model accordingly. Due to it's standardized
message system the containers are testable within standardized tests.

\subsection{Presentational components}\label{presentational-components}

In comparison to the former presentational component is build mostly

template switch

css

shadow dom

\section{5. Thinking further}\label{thinking-further}

Following a \textbf{browsernative first} principle this paper didn't
made use of third-party libraries so far. Contrary to the example in the
last section there are a few reasons why web components shouldn't be
used without any proprietary technology. At first, from a technical
perspective web components are relatively young and at the time of
writing only the newer Chrome and Opera browsers can handle them
correctly. Most probably it will take a long time until native web
components can be safely used without additional legacy support.The
second inevitable question is the notion if they should be used pure
from a \textbf{simplicity perspective} without any proprietary
technology now and future.

Talking about web components it is impossible to avoid technologies like
\textbf{React} altogether. In the last couple years React introduced
some good concepts to web development like the notion of state, passing
data as properties and declarative event management just to name a few.
Many of those design decisions were made to overall \textbf{simplify
browser development through standardized methodologies}. The principles
of microservices could be applied to React development maybe even
further as for example the communication model involves a lot of
delegation and decomposition.

Many of those features and design decisions in React created much of a
hype around the library and technologies like the virtual DOM while web
components never caught up in momentum at a comparable rate. According
to Github Googles Polymer project is even six months older than React
but has less than one third of its stars. Other notable web components
projects like Bosonic\footnote{https://github.com/bosonic/bosonic/} or
Mozilla's X-tag\footnote{https://github.com/x-tag/core} are more dead
than alive, too.

It seems like the web developing community in a broader sense demands
for features and standardization in frameworks like React (and many
others to be fair). Web components and related libraries clearly missed
a selling point in providing standardized ways for developing scalable
components and applications or being to cumbersome to use at all. For
the future it remains unclear if web components will eventually gain
wider adoption within the UI developing community or will be adopted as
low-level technology for framework development emphasized by Sebastian
Markbage, one of the React creators.{[}18{]}

One interesting approach for building web components is the ``reactish''
Skate\footnote{https://github.com/skatejs/skatejs/}. Following the
functional rendering model from React it combines native methods with
additional functionalities known from React like event management. Given
the focus on native technologies it only weights around 4kb (minified
and gzipped) which is ten times less than React. \textbf{Small
encapsulated micro UI frameworks} like Skate could be certainly the near
future of web development and will hopefully gain more attention. The
more mature UI framework Riot already expressed their intention to
gradually develop towards web components.\footnote{http://riotjs.com/compare/\#web-components}

\hypertarget{refs}{}
\hypertarget{ref-Filloux2016}{}
{[}1{]} F. Filloux, ``Bloated html, the best and the worse,'' 2016
{[}Online{]}. Available:
\url{https://mondaynote.com/bloated-html-the-best-and-the-worse-cac6eb06496d}

\hypertarget{ref-Baldwin2006}{}
{[}2{]} C. Y. Baldwin and K. B. Clark, ``Modularity in the Design of
Complex Engineering Systems,'' in \emph{Complex engineered systems:
Science meets technology}, D. Braha, A. A. Minai, and Y. Bar-Yam, Eds.
Berlin, Heidelberg: Springer Berlin Heidelberg, 2006, pp. 175--205
{[}Online{]}. Available:
\href{http://dx.doi.org/10.1007/3-540-32834-3\%7B/_\%7D9}{http://dx.doi.org/10.1007/3-540-32834-3\{\textbackslash{}\_\}9}

\hypertarget{ref-Newman2015}{}
{[}3{]} S. Newman, \emph{Building microservices}. O'Reilly Media, Inc.,
2016.

\hypertarget{ref-Fowler2014}{}
{[}4{]} M. Fowler and J. Lewis, ``Microservices: A definition of this
new architectural term,'' Jan. 2014 {[}Online{]}. Available:
\url{http://www.martinfowler.com/articles/microservices.html}

\hypertarget{ref-Martin}{}
{[}5{]} R. C. Martin, ``The single responsibility principle.''
{[}Online{]}. Available:
\url{http://programmer.97things.oreilly.com/wiki/index.php/The_Single_Responsibility_Principle}

\hypertarget{ref-Kearney2016}{}
{[}6{]} M. Kearney, ``Measure performance with the rail model.'' 2016
{[}Online{]}. Available:
\url{https://developers.google.com/web/fundamentals/performance/rail}

\hypertarget{ref-Conway1968}{}
{[}7{]} M. E. Conway, ``How do committees invent?'' 1968 {[}Online{]}.
Available: \url{http://www.melconway.com/Home/Committees_Paper.html}

\hypertarget{ref-Issa2016}{}
{[}8{]} B. Issa, ``The way of the web.'' Polymer Summit 2016, Oct-2016
{[}Online{]}. Available:
\url{https://www.youtube.com/watch?v=8ZTFEhPBJEE}

\hypertarget{ref-Lawson2016}{}
{[}9{]} N. Lawson, ``Progressive enhancement isn't dead, but it smells
funny.'' 2016 {[}Online{]}. Available:
\url{https://nolanlawson.com/2016/10/13/progressive-enhancement-isnt-dead-but-it-smells-funny/}

\hypertarget{ref-Rota2015}{}
{[}10{]} A. Rota, ``React.js conf 2015 - the complementarity of react
and web components.'' 2015 {[}Online{]}. Available:
\url{https://www.youtube.com/watch?t=124\&v=g0TD0efcwVg}

\hypertarget{ref-Buchner2016}{}
{[}11{]} D. Buchner, ``Demythstifying web components,'' 2016
{[}Online{]}. Available:
\url{http://www.backalleycoder.com/2016/08/26/demythstifying-web-components/}

\hypertarget{ref-vanKesteren2014}{}
{[}12{]} A. van Kesteren, ``Mozilla and web components: Update,'' 2014
{[}Online{]}. Available:
\url{https://hacks.mozilla.org/2014/12/mozilla-and-web-components/}

\hypertarget{ref-Bidelman2016}{}
{[}13{]} E. Bidelman, ``Custom elements v1: reusable web components.''
2016 {[}Online{]}. Available:
\url{https://developers.google.com/web/fundamentals/primers/customelements/}.
{[}Accessed: 01-Dec-2016{]}

\hypertarget{ref-HTML}{}
{[}14{]} \emph{HTML living standard --- last updated 11 january 2017}.
{[}Online{]}. Available: \url{https://html.spec.whatwg.org/multipage/}

\hypertarget{ref-Bidelman2016shadow}{}
{[}15{]} E. Bidelman, ``Shadow dom v1: Self-contained web components.''
2016 {[}Online{]}. Available:
\url{https://developers.google.com/web/fundamentals/getting-started/primers/shadowdom}

\hypertarget{ref-Bidelman2010}{}
{[}16{]} E. Bidelman, ``The basics of web workers.'' 2010 {[}Online{]}.
Available: \url{https://www.html5rocks.com/en/tutorials/workers/basics/}

\hypertarget{ref-Abramov2015}{}
{[}17{]} D. Abramov, ``Presentational and Container Components --
Medium.'' 2015 {[}Online{]}. Available:
\url{https://medium.com/@dan_abramov/smart-and-dumb-components-7ca2f9a7c7d0}.
{[}Accessed: 01-Dec-2016{]}

\hypertarget{ref-Rauschmayer2015}{}
{[}18{]} D. A. Rauschmayer, ``Tree-shaking with webpack 2 and babel 6.''
2015 {[}Online{]}. Available:
\url{http://www.2ality.com/2015/12/webpack-tree-shaking.html}

\end{document}